*Supplementary Material 2*

### 2022-03-25 ms/manuscript.Rmd
Update MS

>>>>

Although some relationships within Polypodiineae (eupolypods I) had previously been resolved differently between various studies using Sanger sequencing, such as Nephrolepidaceae sister to Lomariopsidaceae [@Schuettpelz2007; @Zhang2015] vs. Nephrolepidaceae sister to Tectariaceae, Oleandraceae, Davalliaceae, and Polypodiaceae [@Kuo2011; @Lehtonen2011; @Liu2013; @Testo2016a], our study is in agreement with both nuclear [@Qi2018; @Shen2018] and plastid phylogenomic analyses [@Du2021] that support the latter. Similarly, although Didymochlaenaceae had previously been identified as either sister to the remainder of Polypodiineae [@Kuo2011; @Zhang2015; @Testo2016a] or nested with Hypodematiaceae [@Schuettpelz2007; @Lehtonen2011] by studies using Sanger sequencing, our study as well as nuclear [@Qi2018] and plastid phylogenomic analyses [@Du2021] indicate that Hypodematiaceae is sister to the remainder of the clade.

### 2022-03-25 ms/si.Rmd
Update MS

---
title: "*Supplementary Material 1*"
subtitle: "Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \setlength{\headheight}{13.6pt}
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

tar_load(
  c(
    acc_table_long,
    con_fossil_calibration_tips,
    con_fossil_calibrations_for_treepl,
    fossil_ferns_all,
    plastome_seqs_combined_filtered,
    plastome_tree,
    plastome_tree_rooted,
    ppgi_taxonomy,
    sanger_accessions_selection,
    sanger_sampling,
    sanger_con_tree_dated,
    sanger_con_tree_rooted,
    sanger_alignment_tbl
  ),
  store = ftol_cache
)

@@

# Format references with pandoc
fossils_table <-
fossils_table_raw %>%
  mutate(reference = cite_pd(
    reference_raw, bib = bib, csl = csl, context = temp_context)) %>%
  select(-reference_raw, -notes_on_affinities)

>>>>

---
title: "*Supplementary Material 1*"
subtitle: "Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material 1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \setlength{\headheight}{13.6pt}
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

tar_load(
  c(
    acc_table_long,
    con_fossil_calibration_tips,
    con_fossil_calibrations_for_treepl,
    fossil_ferns_raw,
    fossil_ferns_all,
    plastome_seqs_combined_filtered,
    plastome_tree,
    plastome_tree_rooted,
    ppgi_taxonomy,
    sanger_accessions_selection,
    sanger_sampling,
    sanger_con_tree_dated,
    sanger_con_tree_rooted,
    sanger_alignment_tbl,
    ts_fossil_calibration_tips
  ),
  store = ftol_cache
)

@@

# Format references with pandoc
fossils_table <-
fossils_table_raw %>%
  mutate(reference = cite_pd(
    reference_raw, bib = bib, csl = csl, context = temp_context)) %>%
  select(-reference_raw, -notes_on_affinities) %>%
  # Make sure all references in the table have entries in the bibliography
  assert(function(x) str_detect(x, "\\?", negate = TRUE), reference)

### 2022-03-25 ms/si2.Rmd
Update MS

---
title: "*Supplementary Material 2*"
subtitle: "Fossil Calibration List"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \setlength{\headheight}{13.6pt}
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

@@

tar_load(
  con_fossil_calibration_tips,
  store = ftol_cache
)

fossil_ferns_path <- fs::path(ftol_cache, "user/data_raw/Fossils_Ferns.csv")
```

@@

```{r prep-fossil-data}
# Read all fossils (including those not used in this study)
# Get raw column names
fossils_raw_names <- read_csv(fossil_ferns_path)
# Load data with cleaned column names
fossils_clean_names <- janitor::clean_names(fossils_raw_names)

# Make tibble of fossil data categories
raw_col_names <- colnames(fossils_raw_names)
fossils_categories <-
  tibble(raw_name = raw_col_names) %>%
  mutate(
    category = case_when(
      raw_name %in% raw_col_names[1:4] ~ "Calibration",
      raw_name %in% raw_col_names[5:12] ~ "Fossil Identity",
      raw_name %in% raw_col_names[13:18] ~ "Fossil Age",
      raw_name %in% raw_col_names[19:21] ~ "Fossil Locality",
      raw_name %in% raw_col_names[22:24] ~ "Fossil Specimen",
      TRUE ~ NA_character_
    )
  )

@@

# Tidy fossils in wide format
fossils <-
  fossils_clean_names %>%
  # Filter to only those actually used in the study
  inner_join(
    n_fos_used, by = "n_fos"
  ) %>%
  verify(all(sort(.$n_fos) == sort(n_fos_used$n_fos)))

# Make a tibble of column names for formatting
name_tbl <- tibble(
  raw_name = colnames(fossils_raw_names),
  var = colnames(fossils_clean_names)
) %>%
  right_join(
    tibble(var = colnames(fossils)), by = "var"
  ) %>%
  mutate(var_order = 1:nrow(.))

# Convert to long format
fossils_long <- 
  fossils %>%
  mutate(
    across(everything(), as.character),
    fossil_taxon_header = glue("†{fossil_taxon}")) %>%
  pivot_longer(names_to = "var", values_to = "value", -fossil_taxon) %>%
  left_join(name_tbl, by = "var") %>%
  mutate(
    text = as.character(glue::glue("{raw_name}: {value}"))
  ) %>%
  left_join(fossils_categories, by = "raw_name") %>%
  arrange(fossil_taxon, category, var_order)
```

These data are exported from the FernCal v1.0 database (https://github.com/pteridophy/FossilFerns), which includes `r nrow(fossils_clean_names)` records of fossils to use as calibration points for molecular dating of ferns.

>>>>

---
title: "*Supplementary Material 2*"
subtitle: "Fossil Calibration List"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "frontiers.csl"]
  bookdown::word_document2:
    pandoc_args: [ "--csl", "frontiers.csl"]
    toc: no
    number_sections: no
    reference_docx: Frontiers_Template_SI.docx
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Supplementary Material 2} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \setlength{\headheight}{13.6pt}
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: manual
---

@@

```{r load-functions-si-2, eval = params$doc_type == "manual"}
# Load functions (only if running interactively)
source("_targets.R")
```

@@

tar_load(
  c(
    con_fossil_calibration_tips,
    fossil_ferns_raw,
    fossil_ferns_all),
  store = ftol_cache,
)

# Targets from MS workflow
tar_load(
  fossil_notes,
  store = here::here("_targets")
)
```

@@

```{r tidy-fossil-data}
# fossil_ferns_raw is read in from Fossils_Ferns.csv directly
# fossil_ferns_all is same, but with clean names by janitor::clean_names()

# Make tibble of fossil data categories
raw_col_names <- colnames(fossil_ferns_raw)
fossils_categories <-
  tibble(raw_name = raw_col_names) %>%
  mutate(
    category = case_when(
      raw_name %in% raw_col_names[1:4] ~ "Calibration",
      raw_name %in% raw_col_names[5:12] ~ "Fossil Identity",
      raw_name %in% raw_col_names[13:18] ~ "Fossil Age",
      raw_name %in% raw_col_names[19:21] ~ "Fossil Locality",
      raw_name %in% raw_col_names[22:24] ~ "Fossil Specimen",
      TRUE ~ NA_character_
    )
  )

@@

# Tidy fossils in wide format
fossils <-
  fossil_ferns_all %>%
  # Filter to only those actually used in the study
  inner_join(
    n_fos_used, by = "n_fos"
  ) %>%
  verify(all(sort(.$n_fos) == sort(n_fos_used$n_fos))) %>%
  # Replace with formatted data
  select(
    -node_calibrated, -full_taxon_name, 
    -affinities, -notes_on_affinities, -notes_on_age) %>%
  left_join(fossil_notes, by = "n_fos") %>%
  # Put column names back in proper order
  select(all_of(colnames(fossil_ferns_all)))
```

```{r write-new-formatted, eval = FALSE}
# If needed, write out notes for further editing (if need to add new rows)
fossil_ferns_all %>%
  inner_join(
    n_fos_used, by = "n_fos"
  ) %>%
  anti_join(fossil_notes, by = "n_fos") %>%
  select(n_fos,
         node_calibrated, full_taxon_name, 
         affinities, notes_on_affinities, notes_on_age) %>%
  bind_rows(fossil_notes) %>%
  assert(is_uniq, n_fos) %>%
  arrange(n_fos) %>%
  write_csv(here::here("_targets/user/data_raw/fossils_used_formatted.csv"))
```

```{r format-for-printing}
# Make a tibble of column names for formatting
name_tbl <- tibble(
  raw_name = colnames(fossil_ferns_raw),
  var = colnames(fossil_ferns_all)
) %>%
  right_join(
    tibble(var = colnames(fossils)), by = "var"
  ) %>%
  mutate(var_order = 1:nrow(.))

# Convert to long format
fossils_long <- 
  fossils %>%
  mutate(
    across(everything(), as.character),
    fossil_taxon_header = fossil_taxon) %>%
  pivot_longer(names_to = "var", values_to = "value", -fossil_taxon) %>%
  left_join(name_tbl, by = "var") %>%
  mutate(
    text = as.character(glue::glue("{raw_name}: {value}"))
  ) %>%
  left_join(fossils_categories, by = "raw_name") %>%
  arrange(fossil_taxon, category, var_order) %>%
  # Drop NA entries
  filter(!is.na(value))
```

These data are exported from the FernCal v1.0 database (https://github.com/fernphy/ferncal), which includes `r nrow(fossil_ferns_raw)` records of fossils to use as calibration points for molecular dating of ferns.
